Immutable Source-Mapped Abstract Syntax Tree: \\ A Design Pattern for Refactoring Engine APIs

### 2013-10-01 paper.tex
Fixed references

One option is to provide an API that allows clients to specify what changes
they want to make in terms of AST nodes.  This API can then translate these
into string edits.  For example, an API might offer a \textit{Rewriter} class.
The client would inform the \textit{Rewriter} that it wants to effectively
delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{jdt-astrewrite,cdt-refactoring} take this approach
(with slightly different class and method names).

\begin{figure}[!bt]
\begin{center}
\includegraphics[width=\textwidth]{seq-diagram.eps}
\end{center}
\caption{Sequence diagram illustrating use of a Rewriter to modify source code.}
\label{fig:seq-diagram}
\end{figure}

@@

Both Clang and the original refactoring engine in Xcode~3.0 modify source code
using offset-length information directly.  Notably, Clang offers a
\textit{RewriteRope} class---essentially, a custom string class that allows for
efficient mid-string insertions and deletions~\cite{clang-rewriterope}.
(It uses a B-tree structure inspired by, but not identical to, that described
in the original paper on Ropes~\cite{boehm95ropes}.)

@@

Eclipse JDT provides a class (\textit{ASTRewrite}) that allows the programmer
to specify modifications in terms of Java AST nodes, and then translates those
changes into an LTK \textit{TextEdit} object~\cite{jdt-astrewrite}.  Eclipse
CDT's refactoring engine is based on JDT's and follows a similar design; it
offers its own version of \textit{ASTRewrite}~\cite{cdt-refactoring}.

Refactoring support in the Scala IDE for
Eclipse~\cite{scalaide,scalarefactoring} takes a functional approach,
implementing refactorings as tree transformations.  Source generation is
handled separately: one printer uses position information to retain the
formatting of the original code, while another pretty-prints new nodes that
have been added to the AST~\cite{stocker10scala}.

>>>>

One option is to provide an API that allows clients to specify what changes
they want to make in terms of AST nodes.  This API can then translate these
into string edits.  For example, an API might offer a \textit{Rewriter} class.
The client would inform the \textit{Rewriter} that it wants to effectively
delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{cdt-refactoring} take this approach (with slightly different class
and method names).

\begin{figure}[!bt]
\begin{center}
\includegraphics[width=.9\textwidth]{seq-diagram.eps}
\end{center}
\caption{Sequence diagram illustrating use of a Rewriter to modify source code.}
\label{fig:seq-diagram}
\end{figure}

@@

Both Clang and the original refactoring engine in Xcode~3.0 modify source code
using offset-length information directly.  Notably, Clang offers a
\textit{RewriteRope} class---essentially, a custom string class that allows for
efficient mid-string insertions and
deletions\footnote{\url{http://clang.llvm.org/doxygen/classclang_1_1RewriteRope.html}}.
(It uses a B-tree structure inspired by, but not identical to, that described
in the original paper on Ropes~\cite{boehm95ropes}.)

@@

Eclipse JDT provides a class (\textit{ASTRewrite}\footnote{\url{http://help.eclipse.org/kepler/topic/org.eclipse.jdt.doc.isv/reference/api/org/eclipse/jdt/core/dom/rewrite/ASTRewrite.html}}) that allows the programmer
to specify modifications in terms of Java AST nodes, and then translates those
changes into an LTK \textit{TextEdit} object.  Eclipse CDT's refactoring engine
is based on JDT's and follows a similar design; it offers its own version of
textit{ASTRewrite}~\cite{cdt-refactoring}.

Refactoring support in the Scala IDE for
Eclipse\footnote{\url{http://scala-refactoring.org},
\url{http://scala-ide.org/}} takes a functional approach, implementing
refactorings as tree transformations.  Source generation is handled separately:
one printer uses position information to retain the formatting of the original
code, while another pretty-prints new nodes that were added to the
AST~\cite{stocker10scala}.
